The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-08 manuscript/manuscript.Rnw
use underscore in elephant descriptions

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \elephant are shown in \cref{tab:top50_elephant_placenta}
with a list of all genes in
\cref{tab:all_loxodonta_africana_placenta}. The gene with the
highest expression was the natriuretic peptide precursor B
(\Sexpr{gene.expression("loxodonta africana","NPPB",complete=TRUE)}).
Two thymosins, %
\Sexpr{gene.expression("loxodonta africana","TMSB10",complete=TRUE)} %
and %
\Sexpr{gene.expression("loxodonta africana","TMSB4Y",complete=TRUE)}
were also also was highly expressed. The miRNA mir-675
(\Sexpr{gene.expression("loxodonta
  africana","ENSLAFG00000027406",complete=TRUE)}). which is imbedded
in the first exon of the maternally imprinted long non-coding RNA gene
\textit{H19} \cite{Keniry.ea2012:H19lincRNAisdevelopmental}, a gene
which is currently un-annotated in \elephant{}. %
\DLA{Probably consider fixing the unannotation of H19; it seems pretty
  obvious to me, though it would require fixing the GTF file for
  elephant with the right 
  \href{http://useast.ensembl.org/Loxodonta_africana/Transcript/Summary?db=core;pt=GENSCAN00000034833;r=scaffold_145:855987-899506}{genscan prediction} for this region}
Five uncharacterized genes are expressed to a high degree in the
\elephant placenta (\cref{tab:top50_elephant_placenta}), three of
which %
%
(\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000028493",complete=TRUE)},
\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000032011",complete=TRUE)},
\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000031933",complete=TRUE)}) %
%
are members of an \elephant{}-specific expansion of 11 genes closely
related to the thymus-specific serine protease gene \textit{PRSS16}
\cref{fig:prss16_evolution_tree} The other 8 \textit{PRSS16}-like
genes are also expressed to a lesser degree %
(\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000026505",complete=TRUE)}). %
%
Two other novel genes with high expression are related to the gene
\textit{SSLP1} %
%
(\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000032659",complete=TRUE)} %
and %
%
\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000029551",complete=TRUE)},
%
\cref{fig:sslp1_evolution_tree}), which is also expressed in \cow and
\dog placentas (\cref{tab:top50_dog_placenta,tab:top50_cow_placenta}).
Placental expression of the hormones relaxin (\textit{RLN}, %
\Sexpr{gene.expression("loxodonta africana","ENSLAFG00000036090")})
and prolactin (\textit{PRL}) are exceptionally high in the elephant
when compared to other mammals
(\cref{fig:rln2_evolution_tree,fig:prolactin_evolution_tree}). There
is little to no placental expression of the estrogen synthase gene,
\textit{CYP19A1} %
(\Sexpr{gene.expression("loxodonta africana","CYP19A1")}) and the
enzyme that produces progesterone HSD3B1 %
(\Sexpr{gene.expression("loxodonta africana","HSD3B")}). Other genes
with high expression in the elephant placenta include genes involved
in oxidative stress %
(\Sexpr{gene.expression("loxodonta africana","TXN",complete=TRUE)}),
growth %
(\Sexpr{gene.expression("loxodonta africana","IGF2",complete=TRUE)})
and %
\Sexpr{gene.expression("loxodonta africana","PAEP",complete=TRUE)}, %
a gene usually expressed in the endometrium of the uterus in humans
\cite{Rutanen.ea1987:Progesterone-associatedproteinsPP12and}.

>>>>

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \elephant are shown in \cref{tab:top50_elephant_placenta}
with a list of all genes in
\cref{tab:all_loxodonta_africana_placenta}. The gene with the highest
expression was the natriuretic peptide precursor B
(\Sexpr{gene.expression("loxodonta_africana","NPPB",complete=TRUE)}).
Two thymosins,
\Sexpr{gene.expression("loxodonta_africana","TMSB10",complete=TRUE)}
and
\Sexpr{gene.expression("loxodonta_africana","TMSB4Y",complete=TRUE)}
were also also was highly expressed. The miRNA mir-675
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000027406",complete=TRUE)}).
which is imbedded in the first exon of the maternally imprinted long
non-coding RNA gene \textit{H19}
\cite{Keniry.ea2012:H19lincRNAisdevelopmental}, a gene which is
currently un-annotated in \elephant{}. \DLA{Probably consider fixing
  the unannotation of H19; it seems pretty obvious to me, though it
  would require fixing the GTF file for elephant with the right
  \href{http://useast.ensembl.org/Loxodonta_africana/Transcript/Summary?db=core;pt=GENSCAN00000034833;r=scaffold_145:855987-899506}{genscan
    prediction} for this region} Five uncharacterized genes are
expressed to a high degree in the \elephant placenta
(\cref{tab:top50_elephant_placenta}), three of which
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000028493",complete=TRUE)},
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000032011",complete=TRUE)},
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000031933",complete=TRUE)})
are members of an \elephant{}-specific expansion of 11 genes closely
related to the thymus-specific serine protease gene \textit{PRSS16}
\cref{fig:prss16_evolution_tree} The other 8 \textit{PRSS16}-like
genes are also expressed to a lesser degree
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000026505",complete=TRUE)}).
Two other novel genes with high expression are related to the gene
\textit{SSLP1}
(\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000032659",complete=TRUE)}
and
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000029551",complete=TRUE)},
\cref{fig:sslp1_evolution_tree}), which is also expressed in \cow and
\dog placentas (\cref{tab:top50_dog_placenta,tab:top50_cow_placenta}).
Placental expression of the hormones relaxin (\textit{RLN}, %
\Sexpr{gene.expression("loxodonta_africana","ENSLAFG00000036090")})
and prolactin (\textit{PRL}) are exceptionally high in the elephant
when compared to other mammals
(\cref{fig:rln2_evolution_tree,fig:prolactin_evolution_tree}). There
is little to no placental expression of the estrogen synthase gene,
\textit{CYP19A1}
(\Sexpr{gene.expression("loxodonta_africana","CYP19A1")}) and the
enzyme that produces progesterone HSD3B1
(\Sexpr{gene.expression("loxodonta_africana","HSD3B")}). Other genes
with high expression in the elephant placenta include genes involved
in oxidative stress
(\Sexpr{gene.expression("loxodonta_africana","TXN",complete=TRUE)}),
growth
(\Sexpr{gene.expression("loxodonta_africana","IGF2",complete=TRUE)})
and
\Sexpr{gene.expression("loxodonta_africana","PAEP",complete=TRUE)}, a
gene usually expressed in the endometrium of the uterus in humans
\cite{Rutanen.ea1987:Progesterone-associatedproteinsPP12and}.
